---
title: "GAPMINDER LIFE EXPECTANCIES 2007"
format: dashboard
theme: lux
---

```{r}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, gapminder,
               bslib, bsicons, shiny,
               rnaturalearth, countrycode, plotly,
               htmltools, reactable)
```

```{r}
gap_07 <- gapminder %>%
            filter(year == 2007)
# value_box 1
highest_lide_exp_country <- gap_07 %>%
                              arrange(-lifeExp) %>%
                              head(1) %>%
                              pull(country)

highest_lide_exp_lifeExp <- gap_07 %>%
                              arrange(-lifeExp) %>%
                              head(1) %>%
                              pull(lifeExp)%>%
                              round(1)

# value_box 2
lowest_lide_exp_country <- gap_07 %>%
                              arrange(lifeExp) %>%
                              head(1) %>%
                              pull(country)

lowest_lide_exp_lifeExp <- gap_07 %>%
                              arrange(lifeExp) %>%
                              head(1) %>%
                              pull(lifeExp)%>%
                              round(1)
# value_box 3
average_lide_exp_lifeExp <-
   gap_07 %>%
   summarize(weighted_mean = sum(lifeExp * pop) / sum(pop)) %>%
   round(1)
```


```{r}
country_shapes <- rnaturalearth::ne_countries()

# ggplot(country_shapes) + geom_sf()

gap_07 <- gap_07 %>%
  mutate(country_code = countrycode(country, "country.name", "iso3c"))

# setdiff(gap_07$country_code, country_shapes$adm0_a3)
# "BHR" "COM" "HKG" "MUS" "REU" "STP" "SGP" "PSE"

gap_mapdf <- left_join(country_shapes, gap_07,
            by = c("adm0_a3" = "country_code")) %>%
            mutate(tooltip_label = paste(country, round(lifeExp, 1), sep = ": "))
  
gap_map <-  ggplot(gap_mapdf)+
              geom_sf(aes(fill = lifeExp, text = tooltip_label))+
              theme_void()+
              theme(legend.position = "none")

gap_map_ggplotly <- ggplotly(gap_map, tooltip = "text")
```

```{r}
top_20_countries <- gap_07 %>%
  arrange(desc(lifeExp))%>%
  head(20)%>%
  mutate(tooltip_label = paste(country, round(lifeExp, 1), sep = ": ")) %>%
  ggplot(aes(x=lifeExp,
             y=reorder(country, lifeExp), 
             fill = lifeExp, text = tooltip_label)) +
  geom_col()+
  geom_text(aes(label = round(lifeExp, 1)), nudge_x = -10, color = "white")+
  labs(x= "Life Expectancy", y= "Country")+
  theme(legend.position = "none")

top_20_countries_ggplotly <- ggplotly(top_20_countries, tooltip = "text")
```


# HOME

## Row 1 {height=25%}

```{r}
# value_box1
value_box(
  title = "Highest Life Expectancy",
  value = highest_lide_exp_lifeExp,
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg="#5573b5"),
  p(paste0("(", highest_lide_exp_country, ")"))
)
```

```{r}
# value_box2
value_box(
  title = "Lowest Life Expectancy",
  value = lowest_lide_exp_lifeExp,
  showcase = bsicons::bs_icon("thermometer-half"),
  theme = value_box_theme(bg="#122e6b"),
  p(paste0("(", lowest_lide_exp_country, ")"))
)
```

```{r}
# value_box3
value_box(
  title = "Average Life Expectancy",
  value = average_lide_exp_lifeExp,
  showcase = bsicons::bs_icon("graph-up"),
  theme = value_box_theme(bg="#094bdb")
)
```

## Row 2 {height=75%}

### {width=70%}

```{r title = "MAP of Countries by life Expectancy"}
gap_map_ggplotly
```

### {width=30%}

```{r title = "Top 20 Countries by Life Expectancy"}
top_20_countries_ggplotly
```


# DOWNLOAD DATA

The data used in this dashboard is show and can be downloaded as a CSV.

```{r}

library(htmltools)

htmltools::browsable(
  tagList(
    reactable(
      gapminder,
      elementId = "gapminder-table",
      searchable = T, filterable = T
    ),
     tags$button("Download as CSV", 
                 onclick = "Reactable.downloadDataCSV('gapminder-table')"),
  )
)

```


# ABOUT

This data comes from the r `gapminder` package, and is original sourced from the Gapminder Foundation.

The Gapminder Foundation is a non-profit venture registered in Stockholm, Sweden, that promotes sustainable global development and achievement of the United Nations Millennium Development Goals by increased use and understanding of statistics and other information about social, economic, and environmental development at local, national, and global levels.

Gapminder was founded in 2005 by Ola Rosling, Anna Rosling Rönnlund, and Hans Rosling. The name Gapminder was derived from the "Mind the Gap" warning messages on the London Underground.

An example of one of Hans Rosling's vidéos is show below : 

<iframe width="560" height="315" src="https://www.youtube.com/embed/hVimVzgtD6w?si=HEMI_6MYFhzGeu2t" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

